Towards Pandemic-Scale Ancestral Recombination Graphs of SARS-CoV-2

### 2023-05-18 paper.tex
Yans updates

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/wide_arg_recombination_intervals.pdf}
\caption{\label{fig:breakpoint-distribution}
Distribution of recombination breakpoints and mutations along the genome in
the Wide ARG. Top panel shows the intervals for 1769 breakpoints associated
with 1522 recombination nodes $>1$ descending samples plotted along the genome
as line segments, coloured by interval width. The inset histogram shows the
distribution of these interval widths.
The bottom panel shows the number of intervals that intersect with
each site along the genome (left axis, orange), and the number of mutations
per site (right axis, blue).
See Figure~\ref{fig:long_arg_breakpoint_distribution} for the equivalent plot
for the Long ARG.}
\end{figure}

@@

In total out of the 31 Pango recombinant lineages in the Long ARG we identify 16 as
having a single origin. Supplementary Table~\ref{tab:pango-single-origin} summarises these
together with their inferred parents, showing good agreement between \texttt{sc2ts} and
the official Pango designated parents, with \texttt{sc2ts} providing more a precise
parent designation. Six of the 16 lineages (XA, XAC, XAE, XF, XK, and XS) are of the
simple-origin (XA) type, whereas the rest are of the XAG type, and include multiple Pango
designated lineages as descendants of the originating recombination event. In some cases,
these may be, however, be a result of erroneous Pango designations.

\subsubsection{Multiple origins}
The ARG inference has no pre-defined knowledge
of Pango X lineage assignments, and there is therefore no particular
requirement that all the samples assigned to a given lineage must trace back to
a single recombination event. In this section we examine several Pango X lineage
samples that are inferred to have multiple recombinant origins (i.e. trace
back to independent recombination nodes in the ARG).

>>>>

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/wide_arg_recombination_intervals.pdf}
\caption{\label{fig:breakpoint-distribution}
Distribution of recombination breakpoints and mutations along the genome in
the Wide ARG. Top panel shows the intervals for 1769 breakpoints associated
with 1522 recombination nodes $>1$ descending samples plotted along the genome
as line segments, coloured by interval width. The inset histogram shows the
distribution of these interval widths.
The bottom panel shows the number of intervals that intersect with
each site along the genome (left axis, orange), and the number of mutations
per site (right axis, blue).
TODO describe mutation threshold.
See Figure~\ref{fig:long_arg_breakpoint_distribution} for the equivalent plot
for the Long ARG.}
\end{figure}

@@

Six of the 16 lineages (XA, XAC, XAE, XF, XK, and XS) are of the
simple-origin (XA) type with no other Pango designations among their descendants. 
The remaining 10 are of the XAG type with multiple Pango
designated lineages as additional descendants of the originating recombination event. In some cases,
these may be, however, be a result of erroneous Pango designations. Supplementary Table~\ref{tab:pango-recombinants} also shows the official Pango designated parent lineages for
each recombinant lineage. The parent lineages in the Long ARG agree to a great extent with these, although \texttt{sc2ts} provides more a precise
parent designation.

\subsubsection{Independent multiple origins}
The ARG inference has no pre-defined knowledge
of Pango X lineage assignments, and there is therefore no particular
requirement that all the samples assigned to a given lineage must trace back to
a single recombination event. 15 Pango X lineages are inferred to have multiple
recombinant origins, such that their samples trace to more than one most recent
recombination node in the ARG. Of these, 11 are cases where none of the
recombinants are nested (i.e. there is no overlap between the descendant samples
of each recombination node). In 9 of these 11, there is a single ``main''
recombination event from which the majority of the corresponding recombinant samples
descend. In other two, there are too few Pango X samples to decide on a ``main'' clade.
